Introduction

### 2024-06-21 Presentation/slides2.qmd
full site re-render

---
title: "The Role of Variability in Learning Generalization: A Computational Modeling Approach"
page-layout: full
aliases:
  - /slides.html
css: /Assets/Style/slides.css
format: 
  # html: 
  #   subtitle: "Dissertation Defense"
  # clean-revealjs:
  #   #theme: moon #white
  #   #output-ext: revealjs.html
  #   # template-partials:
  #   #   - assets/title-slide.html
  #   output-file: slides_revealjs
  #   date: 05/28/204
  #   slideNumber: true
  #   hash-type: number # Use slide numbers in the URL hash
  #   center: true
  #   scrollable: true
  #   chalkboard:
  #     theme: whiteboard
  # # pptx:
  #   incremental: false
  # hugo-md:
  #   include: false
  #   html-math-method: mathjax
  #   output-file: slides_hugo.md
  # gfm:
  #   output: false
  #   output-file: slides_gfm.md
  beamer: default
    # incremental: false
    # colortheme: lily 
author:
  - name: Thomas Gorman
    url: https://tegorman13.github.io
    orcid: 0000-0001-5366-5442
    affiliations:
      - name: Indiana University
        department: Psychological & Brain Sciences, Cognitive Science
# params:
#   start_slide: 5
#engine: knitr
---

>>>>

---
title: "The Role of Variability in Learning Generalization: A Computational Modeling Approach"
page-layout: full
# aliases:
#   - /slides.html
# css: /Assets/Style/slides.css
#format: 
  # html: 
  #   subtitle: "Dissertation Defense"
  # clean-revealjs:
  #   #theme: moon #white
  #   #output-ext: revealjs.html
  #   # template-partials:
  #   #   - assets/title-slide.html
  #   output-file: slides_revealjs
  #   date: 05/28/204
  #   slideNumber: true
  #   hash-type: number # Use slide numbers in the URL hash
  #   center: true
  #   scrollable: true
  #   chalkboard:
  #     theme: whiteboard
  # # pptx:
  #   incremental: false
  # hugo-md:
  #   include: false
  #   html-math-method: mathjax
  #   output-file: slides_hugo.md
  # gfm:
  #   output: false
  #   output-file: slides_gfm.md
  #beamer: default
    # incremental: false
    # colortheme: lily 
author:
  - name: Thomas Gorman
    url: https://tegorman13.github.io
    orcid: 0000-0001-5366-5442
    affiliations:
      - name: Indiana University
        department: Psychological & Brain Sciences, Cognitive Science
# params:
#   start_slide: 5
#engine: knitr
---

### 2024-06-21 Sections/Discussion.qmd
full site re-render

---
title: General Discussion
date: today
---

@@

Although theoretically motivated, the IGAS model of Project 1 is best categorized as a descriptive measurement-model. Sufficient to account for group differences, but lacking the machinery necessary to provide a full process-level account of how the empirical quantities of interest are generated. In contrast, Project 2 (HTW) implemented a more robust computational modeling approach, implementing and comparing full process models (ALM & EXAM), capable of generating predictions for both the learning and testing stages of the experiment. ALM and EXAM have been used as models of function learning, cue judgement, and forecasting behavior in numerous studies over the past 25 years [@deloshExtrapolationSineQua1997; @mcdanielPredictingTransferPerformance2009; @brownUnderestimationLinearFunction2017; @kwantesItemOrderMatters2012; @kelleyComparisonModelsLearning2008; @kaneApplicationsBiasVariance2020; @vonhelversenLearningMultiplecueJudgment2010]. The present work presents the first application of these models to to the study of training variability in a visuomotor function learning task. We fit both models to individual participant data, using a form of simulation-based bayesian parameter estimation that allowed us to generate and compare the full posterior predictive distributions of each model. EXAM provided the best overall account of the testing data, and the advantage of EXAM over ALM was significantly greater for the constant group. Notably, EXAM captured the constant groups' ability to extrapolate linearly to novel velocity bands, despite receiving training from only a single input-output pair. This finding suggests that EXAM's linear extrapolation mechanism, combined with the assumption of prior knowledge about the origin point (0, 0), was sufficient to account for the constant groups' accurate extrapolation performance. Such findings may offer a preliminary suggestion that experience with a more variable set of training examples  may be detrimental to performance in simple extrapolation tasks.

### Differences between the two Projects

@@

\ \ \ \

>>>>

---
title: General Discussion
date: today
format:
  html: default
  hugo-md:
    #echo: false
    #output: false
    html-math-method: mathjax
    output-file: dissc_hugo.md
    mermaid-format: png
  gfm:
    #echo: false
    output-file: dissc_gfm.md
    mermaid-format: png
---

# General Discussion

@@

Although theoretically motivated, the IGAS model of Project 1 is best categorized as a descriptive measurement-model. Sufficient to account for group differences, but lacking the machinery necessary to provide a full process-level account of how the empirical quantities of interest are generated. In contrast, Project 2 (HTW) implemented a more robust computational modeling approach, implementing and comparing full process models (ALM & EXAM), capable of generating predictions for both the learning and testing stages of the experiment. ALM and EXAM have been used as models of function learning, cue judgement, and forecasting behavior in numerous studies over the past 25 years [@deloshExtrapolationSineQua1997; @mcdanielPredictingTransferPerformance2009; @brownUnderestimationLinearFunction2017; @kwantesItemOrderMatters2012; @kelleyComparisonModelsLearning2008; @kaneApplicationsBiasVariance2020; @vonhelversenLearningMultiplecueJudgment2010]. The present work presents the first application of these models to to the study of training variability in a visuomotor function learning task. We fit both models to individual participant data, using a form of simulation-based Bayesian parameter estimation that allowed us to generate and compare the full posterior predictive distributions of each model. EXAM provided the best overall account of the testing data, and the advantage of EXAM over ALM was significantly greater for the constant group. Notably, EXAM captured the constant groups' ability to extrapolate linearly to novel velocity bands, despite receiving training from only a single input-output pair. This finding suggests that EXAM's linear extrapolation mechanism, combined with the assumption of prior knowledge about the origin point (0, 0), was sufficient to account for the constant groups' accurate extrapolation performance. Such findings may offer a preliminary suggestion that experience with a more variable set of training examples  may be detrimental to performance in simple extrapolation tasks.

## Differences between the two Projects

@@

\ \

# References
::: {#refs}
:::

### 2024-06-21 Sections/Introduction.qmd
full site re-render

---
title: Introduction
format:
  html: default
  hugo-md:
    echo: false
    output: false
    html-math-method: mathjax
    output-file: d_intro_hugo.md
  gfm:
    echo: true
    output-file: d_into_gfm.md
  hikmah-manuscript-pdf:
    echo: false
    output-file: "intro_pdf.pdf"
    mainfont: "Linux Libertine O"
    mainfontoptions:
      - "Numbers=Proportional"
      - "Numbers=OldStyle"
    mathfont: "Libertinus Math"
---

@@

>>>>

---
title: Introduction
format:
  html: default
  hugo-md:
    echo: false
    output: false
    html-math-method: mathjax
    output-file: d_intro_hugo.md
  gfm:
    echo: true
    output-file: d_into_gfm.md
  # hikmah-manuscript-pdf:
  #   echo: false
  #   output-file: "intro_pdf.pdf"
  #   mainfont: "Linux Libertine O"
  #   mainfontoptions:
  #     - "Numbers=Proportional"
  #     - "Numbers=OldStyle"
  #   mathfont: "Libertinus Math"
---

@@

{{< pagebreak >}}

### 2024-06-21 Sections/htw_exam.qmd
minor changes
full site re-render

---
# title: EXAM Fits and Predictions
date: last-modified
categories: [Simulation, ALM, EXAM, R]
code-fold: true
code-tools: true
execute: 
  warning: false
  eval: true
---

>>>>

---
# title: EXAM Fits and Predictions
# date: last-modified
# categories: [Simulation, ALM, EXAM, R]
code-fold: true
code-tools: true
execute: 
  warning: false
  eval: false
---
